---
title: "03_uniprot_to_GO_method1"
author: "Aidan Coyle"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Aidan Coyle, afcoyle@uw.edu

Originally written 2021/02/02

Roberts Lab, UW-SAFS

Written for analysis of _Hematodinium_ differential gene expression

# Purpose:

This is used to obtain gene ontology (GO) terms from accession IDs.

# Inputs: 

To run this script, you must have the following:

+ A recently-downloaded, uncompressed copy of the Swiss-Prot database, with all GO terms included - available at https://www.uniprot.org/uniprot/

+ A newline-separated file of accession IDs 

# Output:

+ A tab-separated two-column table of accession IDs and their respective GO terms


```{r}
library(tidyverse)
source("hematodinium_analysis_functions.R")
```

# Description of uniprot_to_GO():

This script utilizes a function I created - uniprot_to_GO() - available in hematodinium_analysis_functions.R

Inputs:

accession_path: A newline-separated file of accession IDs

swissprot_path: path to a downloaded uncompressed copy of the SwissProt database with all GO terms included, available at https://www.uniprot.org/uniprot/

output_path: path to output .txt file

```{r}

# Elevated Day 2 vs. Ambient Day 0+2. Individual libraries 
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/elev0_vs_elev2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/elev0_vs_elev2_indiv_All_GOIDs.txt")

# Ambient Day 0+2+17 + Elevated Day 0 + Lowered Day 2 vs. Elevated Day 2
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb0217_elev0_low0_vs_elev2_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb0217_elev0_low0_vs_elev2_All_GOIDs.txt")


# Elevated Day 0 vs. Elevated Day 2. Individual libraries 
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/elev0_vs_elev2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/elev0_vs_elev2_indiv_All_GOIDs.txt")


# Ambient Day 0 vs. Ambient Day 2. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb0_vs_amb2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb0_vs_amb2_indiv_All_GOIDs.txt")

# Ambient Day 0 vs. Ambient Day 17. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb0_vs_amb17_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb0_vs_amb17_indiv_All_GOIDs.txt")

# Ambient Day 2 vs. Ambient Day 17. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb2_vs_amb17_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb2_vs_amb17_indiv_All_GOIDs.txt")

# Ambient Day 2 vs. Elevated Day 2. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb2_vs_elev2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/cbai_hemat_transcriptomev2.0/allgenes_IDs/amb2_vs_elev2_indiv_All_GOIDs.txt")

```
