---
title: "10_2_summary_corr_matrix.Rmd"
author: "Aidan Coyle"
date: "9/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a quick script meant to create a correlation matrix for the summary table we made in the script earlier

Code comes from Laura Spencer

Load packages, install if necessary

```{r}
# Add all required libraries here
list.of.packages <- c("tidyverse", "corrplot", "PerformanceAnalytics")
# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)


# Load all required libraries
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X))
})
```

Next, we'll read in the data and create our correlation matrix

```{r}
# Read in data
summ.dat <- read.csv("../data/indiv_crab_summary.csv")

# Rename first column to "Crab" to make it clearer
summ.dat <- rename(summ.dat, Crab = "ï..Crab")

# Convert as many data columns as possible to numeric
summ.dat$Crab <- as.numeric(as.factor(summ.dat$Crab))
summ.dat$Treatment <- as.numeric(ordered(summ.dat$Treatment, levels = c("Low", "Amb", "Elev")))
summ.dat$cPCR_positive <- as.numeric(as.factor(summ.dat$cPCR_positive))
summ.dat$imm_mature <- as.numeric(as.factor(summ.dat$imm_mature))

# Since we have NAs present for qPCR-SQ_mean_timepoint2, we'd need to drop that column
# To minimize the data loss, we'll create a new column of mean qPCR value
# For Crabs A-F, that means averaging the two columns. For Crabs G-I, that just means taking timepoint1
# To keep it in a spot we're familiar with, we'll just overwrite qPCR_SQ_mean_timepoint1
summ.dat$qPCR_SQ_mean_timepoint1 <- rowMeans(summ.dat[, c("qPCR_SQ_mean_timepoint1", "qPCR_SQ_mean_timepoint2")], na.rm = TRUE)
# We'll also rename the column and remove the timepoint 2 column
summ.dat <- rename(summ.dat, "qPCR_SQ_mean" = "qPCR_SQ_mean_timepoint1")
summ.dat <- select(summ.dat, -qPCR_SQ_mean_timepoint2)

# Remove columns that aren't informative
summ.dat <- select(summ.dat, -c(Crab, Uniq_ID, death, Timepoints))

# Check class of each column
str(summ.dat)

# Perform Shapiro-Wilk test on each column
shapiro.test(summ.dat$Treatment)      # Fails (p = 0.037)
shapiro.test(summ.dat$cPCR_positive)  # Fails (p = 1.6 x 10^-5)
shapiro.test(summ.dat$qPCR_SQ_mean)   # Passes (p = 0.07)
shapiro.test(summ.dat$cbaiv2.0_plasticity)    # Passes (p = 0.1)
shapiro.test(summ.dat$cbaiv4.0_plasticity)    # Passes (p = 0.9)
shapiro.test(summ.dat$hematv1.6_plasticity)   # Fails (p = 0.002)
shapiro.test(summ.dat$cbaiv2.0_pctaligned)    # Passes (p = 0.1)
shapiro.test(summ.dat$cbaiv4.0_pct_aligned)   # Passes (p = 0.2)
shapiro.test(summ.dat$hematv1.6_pct_aligned)  # Passes (p = 0.3)
shapiro.test(summ.dat$cbaiv2.0_transcripts)   # Passes (p = 0.08)
shapiro.test(summ.dat$cbaiv4.0_transcripts)   # Passes (p = 0.9)
shapiro.test(summ.dat$hematv1.6_transcripts)  # Fails (p = 0.01)
shapiro.test(summ.dat$immune_transcripts_cbaiv2.0)  # Passes (p = 0.5)
shapiro.test(summ.dat$immune_transcripts_cbaiv4.0)  # Passes (p = 0.7) 
shapiro.test(summ.dat$immune_transcripts_hematv1.6) # Fails (p = 0.03) 
shapiro.test(summ.dat$carapace_width)         # Passes (p = 0.06)
shapiro.test(summ.dat$shell_cond)             # Fails (p = 1.7 x 10^-5)
shapiro.test(summ.dat$chela_ht)               # Passes (p = 0.4)
shapiro.test(summ.dat$imm_mature)             # Fails (p = 3.2 x 10^-7)
```
Alright, let's sum that up in a prettier way

| Passes normality test                  | Fails normality test                    |
|----------------------------------------|-----------------------------------------|
| qPCR SQ mean                           | Treatment                               |
| cbai v2.0 plasticity                   | cPCR positive                           |
| cbai v4.0 plasticity                   | hemat v1.6 plasticity                   |
| cbai v2.0 percent aligned              | hemat v1.6 number of transcripts        |
| cbai v4.0 percent aligned              | hemat v1.6 number of immune transcripts |
| hemat v1.6 percent aligned             | shell condition                         |
| cbai v2.0 number of transcripts        | maturity                                |
| cbai v4.0 number of transcripts        |                                         |
| cbai v2.0 number of immune transcripts |                                         |
| cbai v4.0 number of immune transcripts |                                         |
| carapace width                         |                                         |
| chela height                           |                                         |

Looks like generally, our cbai metrics pass normality, while those for Hematodinium do not.

Therefore, we're going to do the following:

To compare variables that pass the Shapiro-Wilk test, we'll use the **Pearson correlation test**

To compare variables that don't pass Shapiro-Wilk, we'll use the **Kendall rank correlation test**

To compare variables that don't pass with those that do, we'll use the **Kendall rank correlation test** as well


```{r}
# Split variables into two tables
norm.dat <- summ.dat %>%
  select(c(qPCR_SQ_mean, cbaiv2.0_plasticity, cbaiv4.0_plasticity, cbaiv2.0_pctaligned, 
           cbaiv4.0_pct_aligned, hematv1.6_pct_aligned, cbaiv2.0_transcripts, cbaiv4.0_transcripts,
           immune_transcripts_cbaiv2.0, immune_transcripts_cbaiv4.0, carapace_width, chela_ht))
nonnorm.dat <- summ.dat
nonnorm.dat <- nonnorm.dat[, !summ.dat %in% norm.dat]

#### NORMAL VARIABLES --------------------------------------------

# Compare normal variables with Pearson correlation test
cor(norm.dat, method = "pearson") %>%
  corrplot(tl.cex = .7)
# Make again so we can save as .png
path <- "../output/correlation/normal_vars_dot_plot.png"
png(path)
cor(norm.dat, method = "pearson") %>%
  corrplot(tl.cex = .7)
dev.off()

# Create chart of correlation
chart.Correlation(norm.dat, histogram = F, pch = 5, method = "pearson")
# Make it again so we can save it as a .png
path <- "../output/correlation/normal_vars_corr_chart.png"
png(path)
chart.Correlation(norm.dat, histogram = F, pch = 8, method = "pearson")
dev.off()

#### NON-NORMAL VARIABLES --------------------------------------------

# Compare normal variables with Kendall rank correlation test
cor(nonnorm.dat, method = "kendall") %>%
  corrplot(tl.cex = .7)
# Make again so we can save as .png
path <- "../output/correlation/non_normal_vars_dot_plot.png"
png(path)
cor(nonnorm.dat, method = "kendall") %>%
  corrplot(tl.cex = .7)
dev.off()

# Create chart of correlation
chart.Correlation(nonnorm.dat, histogram = F, pch = 5, method = "kendall")
# Make it again so we can save it as a .png
path <- "../output/correlation/non_normal_vars_corr_chart.png"
png(path)
chart.Correlation(nonnorm.dat, histogram = F, pch = 8, method = "kendall")
dev.off()

#### NORMAL AND NON-NORMAL VARIABLES

# Compare variables with Kendall rank correlation test
cor(summ.dat, method = "kendall") %>%
  corrplot(tl.cex = .7)
# Make it again, so we can save it as a .png
path <- "../output/correlation/all_variables_corr_dot_plot.png"
png(path)
cor(summ.dat, method = "kendall") %>%
  corrplot(tl.cex = .7)
dev.off()

# Create chart of correlation
chart.Correlation(summ.dat, histogram = F, pch = 5, method = "kendall")
# Make it again so we can save it as a .png
path <- "../output/correlation/all_variables_corr_chart.png"
png(path)
chart.Correlation(summ.dat, histogram = F, pch = 8, method = "kendall")
dev.off()



```

