---
title: "11_1_module_GO-MWU_prep_WGCNA_clusters"
author: "Aidan Coyle"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Previously, we ran GO-MWU on a variety of pairwise comparisons for each transcriptome. That's useful, but there's another way to use GO-MWU.

Essentially, you take a module that was deemed significant by WGCNA, and analyze that in the context of all genes that were analyzed by WGCNA.

To do this, you give GO-MWU two inputs.

- A 2-column table of accession IDs and kME values (with the kME values of genes outside the module set to 0). A header line is required, although the contents are irrelevant.

- A tab-delimited 2-column table of accession IDs and GO terms with no header and only 1 line per gene.

This script produces the CSV needed for GO-MWU, which is a 2-column table of accession IDs and log2 fold changes. A header line is required, although the contents are irrelevant.

GO-MWU also requires another input file which we already created - a tab-delimited 2-column table of accession IDs and GO terms with no header and only 1 line per gene

In this script, we'll work to create the first of those two files.

First, here's a list of all significant modules for all 3 transcriptomes (that aren't associated with a crab). Ones that seem particularly interesting are bolded. Names correspond to what they're called in the hemat_level analysis for that respective transcriptome/comparison combination


| Transcriptome | Comparison | Trait               | Module name | p-value | Lowest?                  | Others                                         |
|---------------|------------|---------------------|-------------|---------|--------------------------|------------------------------------------------|
| Unfiltered    | All        | **day**             | **black**   | 0.04    | Y                        |                                                |
| Unfiltered    | All        | Temp (Low vs Amb)   | tan         | 0.05    | Y                        |                                                |
| Unfiltered    | All        | Temp (Low vs Amb)   | **cyan**    | 0.02    | N (Crab C, 0.002)        | Crab B, Temp (Elev), Hemat_Level               |
| Unfiltered    | All        | Temp (Low vs Elev)  | **brown**   | 0.02    | Y                        | Temp (Elev)                                    |
| Unfiltered    | All        | Temp (Elev)         | **cyan**    | 0.04    | N (Crab C, 0.002)        | Crab B, Temp (Low vs Amb), Hemat_Level         |
| Unfiltered    | All        | Temp (Elev)         | **brown**   | 0.03    | N (Low vs Elev, 0.02)    |                                                |
| Unfiltered    | All        | infection           | red         | 0.05    | N (Crab F, 0.004)        |                                                |
| Unfiltered    | All        | hemat_level         | magenta     | 0.02    | N (Crab C, 6x10^-13)     | CW                                             |
| Unfiltered    | All        | **hemat_level**     | **cyan**    | 0.008   | N (Crab C, 0.002)        | Crab B, Temp (Low vs Amb), Temp (Elev) |
| Unfiltered    | All        | SC                  | red         | 0.05    | N (Crab F, 0.004)        | Infection                                      |
| Unfiltered    | Amb + Elev | hemat_level         | lightcyan   | 0.03    | N (Crab A, 8x10^-4)      |                                                |
| Unfiltered    | Amb + Elev | CW                  | darkgrey    | 0.04    | Tie (Crab C)             |                                                |
| Unfiltered    | Amb + Elev | CW                  | darkorange  | 0.05    | N (Crab H, 3x10^-9)      |                                                |
| Unfiltered    | Amb + Elev | CW                  | skyblue     | 3x10^-4 | N (Crab G, 8x10&-9)      |                                                |
| Unfiltered    | Amb + Elev | CW                  | lightgreen  | 0.03    | N (Crab C, 8x10^-8)      |                                                |
| Host          | All        |**Temp (Low v Amb)** | black       | 6x10^-4 | Y                        | Elev                                           |
| Host          | All        | Temp (Low vs. Amb)  | red         | 0.01    | N (Crab C, 2x10^-8)      | hemat_level, CW                                |
| Host          | All        |**Temp (Low v Elev)**| blue        | 0.02    | Y                        | Elev                                           |
| Host          | All        | Temp (Elev)         | blue        | 0.03    | N (Low vs. Elev, 0.02)   |                                                |
| Host          | All        | Temp (Elev)         | black       | 0.05    | N (Low vs. Amb, 6x10^-4) |                                                |
| Host          | All        | hemat_level         | red         | 0.01    | N (Crab C, 2x10^-8)      | Low vs. Amb, CW                                |
| Host          | All        | CW                  | red         | 0.04    | N (Crab C, 2x10^-8)      | Low vs. Amb, hemat_level                       |
| Host          | All        | SC                  | pink        | 0.03    | N (Crab F, 0.002)        |                                                |
| Host          | Amb + Elev | CW                  | pink        | 0.02    | N (Crab C, 9x10^-6)      |                                                |
| Parasite      | All        | **day**             | black       | 0.04    | Y                        |                                                |
| Parasite      | All        | Temp (Low vs. Amb)  | turquoise   | 0.02    | N (Crab C, 4x10^-6)      | hemat_level, CW                                |
| Parasite      | All        | hemat_level         | turquoise   | 0.01    | N (Crab C, 4x10^-6)      | Low vs Amb, CW                                 |
| Parasite      | All        | **hemat_level**     | blue        | 0.01    | N (Crab B, 0.002)        | Crab B, Crab C                                 |

Modules

**Unfiltered, All**
- Black (day)
- Cyan (temperature, hemat level)
- Brown (temperature)

**Host, All**
- Black (temperature)
- Blue (temperature)

**Parasite, All**
- Black (day)
- Blue (hemat_level)


Alright, let's get started:

First, load all libraries as usual

```{r libraries, message = FALSE, warning=FALSE}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- "tidyverse"
# Add all libraries that are installed using BiocManager here
bioconductor.packages <- ""

# Install BiocManager if needed
if(!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
new.bioc.packages <- bioconductor.packages[!(bioconductor.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)
if(length(new.bioc.packages)) BiocManager::install(new.bioc.packages)

# Load all required libraries
all.packages <- c(list.of.packages, bioconductor.packages)
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})

source("hematodinium_analysis_functions.R")

```


We'll now repeatedly run WGCNA_modules_accessions_KMEs()

This works as follows:


Inputs:
- The transcripts within a module produced by WGCNA
- All kMEs for all modules in your WGCNA run (should be in same directory as significant module)
- A BLAST file for your relevant transcriptome


Output:
  - A tab-separated two-column table of accession IDs and module kME values for all genes (kME values for genes not in module should be set to 0)
      - This goes in the output_GOMWU directory
  - A single column of accession IDs (this will be used to create the second input for GO-MWU)
      - This will be put in the output_accessions directory
      
Function arguments:
- module_name: color of the module you want to examine
- transcriptome: name of the relevant transcriptome
- compar: libraries used in WGCNA. Should be all_crabs_no_filter or amb_vs_elev_no_filter
- blast_path: path to the BLAST file for your relevant transcriptome
- output_GOMWU: path to the folder you want to put your files in (should be the one you run GO-MWU in)
- output_accessions: path to the folder you want to put your single column of accession IDs in

```{r inputs}

# UNFILTERED TRANSCRIPTOME ------------------------------

# Cyan module, cbai2.0, all crabs
WGCNA_modules_accessions_kMEs(module_name = "cyan",
                              transcriptome = "cbai_transcriptomev2.0",
                              compar = "all_crabs_no_filter",
                              blast_path = "../data/cbai_hemat_diamond_blastx_table_transcriptome_v2.0.txt",
                              output_GOMWU = "../scripts/11_4_running_GOMWU/",
                              output_accessions = "../output/accession_n_GOids/WGCNA_modules/")

# Brown module, cbai2.0, all crabs
WGCNA_modules_accessions_kMEs(module_name = "brown",
                              transcriptome = "cbai_transcriptomev2.0",
                              compar = "all_crabs_no_filter",
                              blast_path = "../data/cbai_hemat_diamond_blastx_table_transcriptome_v2.0.txt",
                              output_GOMWU = "../scripts/11_4_running_GOMWU/",
                              output_accessions = "../output/accession_n_GOids/WGCNA_modules/")

#### HOST TRANSCRIPTOME -------------------------------

# Black module, cbai4.0, all crabs
WGCNA_modules_accessions_kMEs(module_name = "black",
                              transcriptome = "cbai_transcriptomev4.0",
                              compar = "all_crabs_no_filter",
                              blast_path = "../data/cbai_blastx_table_transcriptome_v4.0.txt",
                              output_GOMWU = "../scripts/11_4_running_GOMWU/",
                              output_accessions = "../output/accession_n_GOids/WGCNA_modules/")

# Blue module, cbai4.0, all crabs
WGCNA_modules_accessions_kMEs(module_name = "blue",
                              transcriptome = "cbai_transcriptomev4.0",
                              compar = "all_crabs_no_filter",
                              blast_path = "../data/cbai_blastx_table_transcriptome_v4.0.txt",
                              output_GOMWU = "../scripts/11_4_running_GOMWU/",
                              output_accessions = "../output/accession_n_GOids/WGCNA_modules/")

#### PARASITE TRANSCRIPTOME -------------------------

# Black module, hemat1.6, all crabs
WGCNA_modules_accessions_kMEs(module_name = "black",
                              transcriptome = "hemat_transcriptomev1.6",
                              compar = "all_crabs_no_filter",
                              blast_path = "../data/hemat_blastx_table_transcriptome_v1.6.txt",
                              output_GOMWU = "../scripts/11_4_running_GOMWU/",
                              output_accessions = "../output/accession_n_GOids/WGCNA_modules/")

# Blue module, hemat1.6, all crabs
WGCNA_modules_accessions_kMEs(module_name = "blue",
                              transcriptome = "hemat_transcriptomev1.6",
                              compar = "all_crabs_no_filter",
                              blast_path = "../data/hemat_blastx_table_transcriptome_v1.6.txt",
                              output_GOMWU = "../scripts/11_4_running_GOMWU/",
                              output_accessions = "../output/accession_n_GOids/WGCNA_modules/")
```
