---
title: "11_2_prod_cluster_GOterms"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In the previous script, we produced a 2-column table of accession IDs and kME values for each module we wanted to examine.

We also produced a single column of accession IDs in that same script.

Here, we will use that single column to produce the other input necessary for GO-MWU - a **2-column table of accession IDs and GO terms with no header**

We will later need to reduce it to just one line per unique accession ID, but that's simple enough to do later

First, load libraries and functions

```{r libraries, message = FALSE, warning=FALSE}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- "tidyverse"
# Add all libraries that are installed using BiocManager here
bioconductor.packages <- ""

# Install BiocManager if needed
if(!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
new.bioc.packages <- bioconductor.packages[!(bioconductor.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)
if(length(new.bioc.packages)) BiocManager::install(new.bioc.packages)

# Load all required libraries
all.packages <- c(list.of.packages, bioconductor.packages)
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})

# Functions are defined in hematodinium_analysis_functions.R
source("hematodinium_analysis_functions.R")
```

We'll use the same function we've used in earlier runs when trying to create inputs for GO-MWU - uniprot_to_GO()

# Description of uniprot_to_GO():
Inputs:

accession_path: A newline-separated file of accession IDs

swissprot_path: path to a downloaded uncompressed copy of the SwissProt database with all GO terms included, available at https://www.uniprot.org/uniprot/

output_path: path to output .txt file

## All Samples

This WGCNA run examined samples from all crabs to look at correlation to a variety of variables

```{r}
# COMPLETE TRANSCRIPTOME ------------------------------

# Blue module, cbai2.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_blue_GeneIDs.txt ",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_blue_GOIDs.txt")

# Cyan module, cbai2.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_cyan_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_cyan_GOIDs.txt")

# Salmon module, cbai2.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_salmon_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_salmon_GOIDs.txt")

# Yellow module, cbai2.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_yellow_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/all_crabs_yellow_GOIDs.txt")

#### TANNER CRAB TRANSCRIPTOME -------------------------------

# Brown module, cbai4.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev4.0/all_crabs_black_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev4.0/all_crabs_black_GOIDs.txt")

# Black module, cbai4.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev4.0/all_crabs_brown_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev4.0/all_crabs_brown_GOIDs.txt")

#### PARASITE TRANSCRIPTOME -------------------------

# Brown module, hemat1.6, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/hemat_transcriptomev1.6/all_crabs_brown_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/hemat_transcriptomev1.6/all_crabs_brown_GOIDs.txt")

# Pink module, hemat1.6, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/hemat_transcriptomev1.6/all_crabs_pink_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/hemat_transcriptomev1.6/all_crabs_pink_GOIDs.txt")
```

## Ambient and Lowered-Temp Samples

This WGCNA run looked to analyze correlation to change in _Hematodinium_ level. It used samples from all crabs where two qPCR timepoints were available. Due to the elevated-treatment mortality event, that means it only included crabs A-F (the Ambient and Lowered treatment groups).

```{r}
# COMPLETE TRANSCRIPTOME ------------------------------

# Blue module, cbai2.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/amb_low_blue_GeneIDs.txt ",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev2.0/amb_low_blue_GOIDs.txt")

#### TANNER CRAB TRANSCRIPTOME -------------------------------

# Red module, cbai4.0, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev4.0/amb_low_red_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/cbai_transcriptomev4.0/amb_low_red_GOIDs.txt")

#### PARASITE TRANSCRIPTOME -------------------------

# Green module, hemat1.6, all crabs
uniprot_to_GO(accession_path = "../output/accession_n_GOids/WGCNA_modules/hemat_transcriptomev1.6/amb_low_green_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/WGCNA_modules/hemat_transcriptomev1.6/amb_low_green_GOIDs.txt")
```

