---
title: "12_1_supp_file_creation"
author: "Aidan Coyle"
date: "12/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supplemental File Creation

In this script, I will create the supplemental files for the paper.

Since the numbering hasn't been nailed down yet, I'll name the sections by describing the table I'm creating, rather than with numbering.

## Load libraries

```{r}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- "tidyverse"

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)

# Load all required libraries
all.packages <- list.of.packages
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})
```

## Annotation Tables

I'll create an annotation file for each transcriptome. This will be done by merging the existing annotation table with the GO terms for that annotation.

Annotated tables and the GO term database have been downloaded in previous scripts, but for the sake of reproducibility, here are the links to each:

**Transcriptome Annotation Tables**

All are in outfmt 6 and were annotated with BLASTx

[Complete transcriptome, AKA cbai_transcriptome_v2.0](https://gannet.fish.washington.edu/Atumefaciens/20200508_cbai_diamond_blastx_transcriptome-v2.0/20200507.C_bairdi.Trinity.blastx.outfmt6). MD5 is ace82a75cb947574ac807d868427253c

[Tanner crab transcriptome, AKA cbai_transcriptome_v4.0](https://gannet.fish.washington.edu/Atumefaciens/20210318_cbai_diamond_blastx_transcriptome-v4.0/cbai_transcriptome_v4.0.blastx.outfmt6). MD5 is 8fd2ab9c27e59653fcb4f32574be1f61

[Hematodinium transcriptome, AKA hemat_transcriptomev1.6](https://gannet.fish.washington.edu/Atumefaciens/20200814_hemat_diamond_blastx_v1.6_v1.7_v2.1_v3.1/hemat_transcriptome_v1.6.fasta.blastx.outfmt6). MD5 is 698513d787f92a65debb9f70dbb2e32f

**GO terms**

To get GO terms, we downloaded the full Swiss-ProtKB/Uniprot database (which includes all GO terms) on 2021-02-09 from https://www.uniprot.org/uniprot/
MD5sum for that download was ace82a75cb947574ac807d868427253c

**Build annotation tables**
```{r}
#### Setup ----------------
# These will be used to build each full annotation table

# Set up header for outfmt 6
table_head <- c("query_seqid", "subj_seqid", "pident",
                "length", "mismatch", "gapopen",
                "qstart", "qend", "sstart",
                "send", "evalue", "bitscore")

# Read in full Swiss-ProtKB/Uniprot database
sprot_db <- read.delim(file = "../data/all_uniprot_info_inc_GOterms.tab",
                       header = TRUE)

# Rename:
#   Entry column to Accession_ID 
#   length to sprot.length 
sprot_db <- sprot_db %>%
  dplyr::rename(Accession_ID = Entry,
                sprot.length = Length)

#### Complete Transcriptome ---------------------------------

# Read in data
annotation <- read.delim(file = "../data/blast_tables/cbai_transcriptomev2.0_blast_table.txt",
                         header = FALSE,
                         col.names = table_head)

# Second column contains several relevant pieces of info
# Split to extract relevant info (accession ID), which we'll use to merge
# and then remove the "sp" column, which just reads "sp" for all entries
annotation <- annotation %>%
  separate(col = subj_seqid,
           into = c("sp", "Accession_ID", "species"), 
           sep = "\\|") %>%
  select(-sp)

# Left join annotation table with Swiss-Prot database by accession ID
full_annot <- left_join(annotation, sprot_db, by = "Accession_ID")

# Write table to file
write.csv(full_annot, file = "../paper/supp_files/annotation_tables/complete_transcriptome.csv",
          row.names = FALSE)

#### Tanner Crab Transcriptome ---------------------------------

# Read in data
annotation <- read.delim(file = "../data/blast_tables/cbai_transcriptomev4.0_blast_table.txt",
                         header = FALSE,
                         col.names = table_head)

# Second column contains several relevant pieces of info
# Split to extract relevant info (accession ID), which we'll use to merge
# and then remove the "sp" column, which just reads "sp" for all entries
annotation <- annotation %>%
  separate(col = subj_seqid,
           into = c("sp", "Accession_ID", "species"), 
           sep = "\\|") %>%
  select(-sp)

# Left join annotation table with Swiss-Prot database by accession ID
full_annot <- left_join(annotation, sprot_db, by = "Accession_ID")

# Write table to file
write.csv(full_annot, file = "../paper/supp_files/annotation_tables/Tanner_crab_transcriptome.csv",
          row.names = FALSE)

#### Hematodinium Transcriptome ---------------------------------

# Read in data
annotation <- read.delim(file = "../data/blast_tables/hemat_transcriptomev1.6_blast_table.txt",
                         header = FALSE,
                         col.names = table_head)

# Second column contains several relevant pieces of info
# Split to extract relevant info (accession ID), which we'll use to merge
# and then remove the "sp" column, which just reads "sp" for all entries
annotation <- annotation %>%
  separate(col = subj_seqid,
           into = c("sp", "Accession_ID", "species"), 
           sep = "\\|") %>%
  select(-sp)

# Left join annotation table with Swiss-Prot database by accession ID
full_annot <- left_join(annotation, sprot_db, by = "Accession_ID")

# Write table to file
write.csv(full_annot, file = "../paper/supp_files/annotation_tables/Hematodinium_transcriptome.csv",
          row.names = FALSE)
```




