---
title: "35_GO-MWU_prep"
author: "Aidan Coyle"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Aidan Coyle, afcoyle@uw.edu

Originally written 2021/03/10

Roberts Lab, UW-SAFS

Written for analysis of _Hematodinium_ differential gene expression

# Purpose:

This script produces the CSV needed for GO-MWU, which is a 2-column table of accession IDs and unadjusted p-values. A header line is required, although the contents are irrelevant.

GO-MWU also requires another input file which we already created - a tab-delimited 2-column table of accession IDs and GO terms with no header and only 1 line per gene

# Explanation of geneIDs_pvals():

This script utilizes a function built for this analysis by me - geneids_pvals(). It is defined in hematodinium_analysis_functions.R

### Inputs and Outputs

input_file: DESeq2 output file containing transcript IDs and unadjusted p-values

blast_file: path that leads to transcript ID/accession ID table. Optimally, use an annotated BLASTx table of the same transcriptome used to create the kallisto index

output_file: path to the output file - a 2-column CSV of accession IDs and unadjusted p-values with a header line. This should be written to the same directory that you will run GO-MWU in.

```{r pkgsnfunctions, message=FALSE}
library(tidyverse)

source("hematodinium_analysis_functions.R")

```

# Obtaining CSVs

```{r geneIDs_pvals} 
# Elevated Day 0 vs. Elevated Day 2, indiv. libraries only
geneIDs_pvals(input_file =  "../graphs/DESeq2_output/hemat_transcriptomev1.6/elev0_vs_elev2_indiv/AllGenes_wcols.txt",
              blast_file =  "../output/BLASTs/uniprot_swissprot/hemat1.6_blastxres.tab",
              output_file  =  "../scripts/36_running_GO-MWU/hemat1.6_elev0_vs_elev2_indiv_pvals.csv")

# Ambient Day 2 vs. Elevated Day 2, indiv. libraries only
geneIDs_pvals(input_file = "../graphs/DESeq2_output/hemat_transcriptomev1.6/amb2_vs_elev2_indiv/AllGenes_wcols.txt",
              blast_file = "../output/BLASTs/uniprot_swissprot/hemat1.6_blastxres.tab",
              output_file = "../scripts/36_running_GO-MWU/hemat1.6_amb2_vs_elev2_indiv_pvals.csv")
```