---
title: "33_uniprot_to_GO_method1"
author: "Aidan Coyle"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Aidan Coyle, afcoyle@uw.edu

Originally written 2021/03/10
Roberts Lab, UW-SAFS

Written for analysis of _Hematodinium_ differential gene expression

# Purpose:

This is used to obtain gene ontology (GO) terms from accession IDs. 

# Inputs: 

To run this script, you must have the following:

+ A recently-downloaded, uncompressed copy of the Swiss-Prot database, with all GO terms included - available at https://www.uniprot.org/uniprot/

+ A newline-separated file of accession IDs 

# Output:

+ A tab-separated two-column table of accession IDs and their respective GO terms


```{r libraries, message = FALSE, warning=FALSE}
# Add all required libraries here
list.of.packages <- "tidyverse"
# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)


# Load all required libraries
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X))
})

# Functions are defined in hematodinium_analysis_functions.R
source("hematodinium_analysis_functions.R")
```

# Description of uniprot_to_GO():

This script utilizes a function I created - uniprot_to_GO() - available in hematodinium_analysis_functions.R

Inputs:

accession_path: A newline-separated file of accession IDs

swissprot_path: path to a downloaded uncompressed copy of the SwissProt database with all GO terms included, available at https://www.uniprot.org/uniprot/

output_path: path to output .txt file

```{r}

# Elevated Day 0 vs. Elevated Day 2. Individual libraries 
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/elev0_vs_elev2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab", 
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/elev0_vs_elev2_indiv_All_GOIDs.txt")

# Ambient Day 2 vs. Elevated Day 2. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/amb2_vs_elev2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/amb2_vs_elev2_indiv_All_GOIDs.txt")

# Ambient Day 0 vs. Ambient Day 2. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/amb0_vs_amb2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/amb0_vs_amb2_indiv_All_GOIDs.txt")

# Ambient Day 0 vs. Ambient Day 17. Individual libraries
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/amb0_vs_amb17_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/amb0_vs_amb17_indiv_All_GOIDs.txt")

# Lowered Day 0 vs. Lowered Day 2
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_low2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_low2_indiv_All_GOIDs.txt")

# Lowered Day 0 vs. Lowered Day 2 + 17
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_low217_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_low217_indiv_All_GOIDs.txt")

# Lowered Day 0 + Ambient Day 0 + 2 + 17 vs. Lowered Day 2 + 17
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_amb0217_vs_low217_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_amb0217_vs_low217_indiv_All_GOIDs.txt")

# Lowered Day 2 + 17 vs. Ambient Day 2+17
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low217_vs_amb217_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low217_vs_amb217_indiv_All_GOIDs.txt")

# Lowered Day 0 vs. Ambient Day 0
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_amb0_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_amb0_indiv_All_GOIDs.txt")

# Lowered Day 0 vs. Lowered Day 17
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_low17_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_vs_low17_indiv_All_GOIDs.txt")

# Lowered Day 0 + Ambient Day 0 vs. Lowered Day 17 + Ambient Day 17
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_amb0_vs_low17_amb17_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_amb0_vs_low17_amb17_indiv_All_GOIDs.txt")

# Lowered Day 0 + Elevated Day 0 vs. Lowered Day 2 + Elevated Day 2
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_elev0_vs_low2_elev2_indiv_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/low0_elev0_vs_low2_elev2_indiv_All_GOIDs.txt")

# All Genes
uniprot_to_GO(accession_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/all_indiv_libraries_All_GeneIDs.txt",
              swissprot_path = "../data/all_uniprot_info_inc_GOterms.tab",
              output_path = "../output/accession_n_GOids/hemat_transcriptomev1.6/allgenes_IDs/all_indiv_libraries_All_GOIDs.txt")


```
