---
title: "5_14_WGCNA_hemat1.6_AllCrabs"
author: "Aidan Coyle"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Analysis of Ambient and Elevated-Temperature Gene Expression using WGCNA

We will now run WGCNA on all individual libraries 

This will be ran on libraries aligned to hemat_transcriptomev1.6, which only includes genes that are presumably _Hematodinium_.

Note the additional column compared to previous runs - since we are including crabs D and F, which are uninfected, infection status is now an included variable

Table of crabs and libraries included in analysis:
| Crab ID | Treatment Group | Day 0 Sample ID | Day 2 Sample ID | Day 17 Sample ID |Infection Status |
|---------|-----------------|-----------------|-----------------|------------------|-----------------|
| A       | ambient         | 178             | 359             | 463              | infected        |  
| B       | ambient         | 118             | 349             | 481              | infected        | 
| C       | ambient         | 132             | 334             | 485              | infected        | 
| E       | decreased       | 151             | 254             | 445              | infected        | 
| G       | elevated        | 173             | 272             | NA               | infected        | 
| H       | elevated        | 072             | 294             | NA               | infected        | 
| I       | elevated        | 127             | 280             | NA               | infected        | 

Again, script is based largely on [Yaamini's script](https://github.com/eimd-2019/project-EWD-transcriptomics/blob/master/analyses/WGCNA/WGCNA.md), which is based largely on [the official WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/)

We will first extract the TPM (transcripts per million) counts from the kallisto libraries created earlier in the pipeline. We will then change those to logTPM counts, and then begin the WGCNA analysis

All kallisto libraries should be available within the GitHub repo. However, one datafile - our blastx table, hemat_blastx_table_transcriptome_v1.6.txt,is not. This is a BLASTx annotation of hemat_transcriptome_v1.6.fasta.

Annotation description available [here](https://robertslab.github.io/sams-notebook/2020/08/14/Transcriptome-Annotation-Hematodinium-Transcriptomes-v1.6-v1.7-v2.1-v3.1-with-DIAMOND-BLASTx-on-Mox.html), [direct file available here](https://gannet.fish.washington.edu/Atumefaciens/20200814_hemat_diamond_blastx_v1.6_v1.7_v2.1_v3.1/hemat_transcriptome_v1.6.fasta.blastx.outfmt6), and [original transcriptome available here](https://gannet.fish.washington.edu/Atumefaciens/20210308_hemat_trinity_v1.6_v1.7/hemat_transcriptome_v1.6.fasta_trinity_out_dir/hemat_transcriptome_v1.6.fasta), 
md5sum: 698513d787f92a65debb9f70dbb2e32f

Note that for this, minimum module size was reduced to 10, due to the much smaller number of Hematodinium genes


```{r libraries, message = FALSE, warning=FALSE}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- "tidyverse"
# Add all libraries that are installed using BiocManager here
bioconductor.packages <- c("DESeq2", "WGCNA")

# Install BiocManager if needed
if(!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
new.bioc.packages <- bioconductor.packages[!(bioconductor.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)
if(length(new.bioc.packages)) BiocManager::install(new.bioc.packages)

# Load all required libraries
all.packages <- c(list.of.packages, bioconductor.packages)
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})
```

### Step 1: Extract TPM Counts from Kallisto Libraries

This portion of the script is based largely off 21_obtaining_TPM_for_DEGs.Rmd

First, set all variables.

Later in the script, you will need to specify a soft thresholding power to choose. But otherwise, minimal changes are necessary

```{r TPMextract}
# Path to kallisto libraries
kallisto_path <- "../output/kallisto_libraries/hemat_transcriptomev1.6/"

# Libraries we want to read in to our TPM matrix
libraries <- c("178", "118", "132", "151", "173", "072", "127", "359", "349", "334", "254", "272", "294", "280", "463", "481", "485", "445")

# For each row, crab and day should correspond to the order of libraries (ex: 4th row of crabTraits should match libraries[4])

# If treating elevated day 0 as elevated and lowered day 0 as lowered
crabTraits <- data.frame("crab" = c(rep(c("A", "B", "C", "E", "G", "H", "I"), times = 2), 
                                    "A", "B", "C", "E"),
                         "day" = factor(c(rep(0, times = 7),
                                   rep(2, times = 7),
                                   rep(17, times = 4))),
                         "temp" = c(rep(c(rep("Ambient", times = 3),
                                          "Lowered", 
                                          rep("Elevated", times = 3)), times = 2),
                                    rep("Ambient", times = 3),
                                    "Lowered"))

# Use the above one if categorizing Day 0 Elevated-temperature crab as "Elevated" and Day 0 Lowered-temperature crab as "Lowered". If categorizing them as "Ambient" (since they were at ambient temperature when Day 0 samples were taken), use the one below. Comment out the one not used!

# If treating elevated day 0 and lowered day 0 as ambient
crabTraits <- data.frame("crab" = c(rep(c("A", "B", "C", "E", "G", "H", "I"), times = 2), 
                                    "A", "B", "C", "E"),
                         "day" = factor(c(rep(0, times = 7),
                                   rep(2, times = 7),
                                   rep(17, times = 4))),
                         "temp" = c(rep("Ambient", times = 10),
                                    "Lowered",
                                    rep("Elevated", times = 3),
                                    rep("Ambient", times = 3),
                                    "Lowered"))

# Create clinical data trait matrix. Same rules as above, but all columns are numeric are numeric. Crab A will be noted as 1, B as 2, and C as 3. Again, comment out the one not used!

# If treating elevated day 0 as elevated
crabClinicalData <- data.frame("crab" = c(rep(c(1, 2, 3, 4, 5, 6, 7), times = 2), 
                                    1, 2, 3, 4),
                         "day" = c(rep(0, times = 7),
                                   rep(2, times = 7),
                                   rep(17, times = 4)),
                         "temp" = c(rep(c(rep(2, times = 3),
                                          1, 
                                          rep(3, times = 3)), times = 2),
                                    rep(2, times = 3),
                                    1))

# If treating elevated day 0 and lowered day 0 as ambient
crabClinicalData <- data.frame("crab" = c(rep(c(1, 2, 3, 4, 5, 6, 7), times = 2), 
                                    1, 2, 3, 4),
                         "day" = c(rep(0, times = 7),
                                   rep(2, times = 7),
                                   rep(17, times = 4)),
                         "temp" = c(rep(2, times = 10),
                                    1,
                                    rep(3, times = 3),
                                    rep(2, times = 3),
                                    1))

# Variable being examined - should match column in two data frames above
variable <- "temp"

# Start and ending we want for each file and graph saved
file_start <- "../output/WGCNA_output/hemat_transcriptome_v1.6/all_crabs_no_filter/day0_elev_as_amb/min_module_10/"

# Location of blastx table
blastx_table_site <- "../data/hemat_blastx_table_transcriptome_v1.6.txt"

# Number of samples we're examining
numsamples <- 18

# Minimum module size (number of genes per module)
minimum_module <- 10
```

Then, we begin creating our TPM matrix for all transcripts

```{r matrixcreation}
# Create character vector with all filenames for our libraries
kallisto_files <- paste0(kallisto_path, "id", libraries, "/abundance.tsv")

# Read first kallisto file in to start data frame
TPMcounts <- read.delim(file = kallisto_files[1],
                        header = TRUE,
                        sep = "\t")
# Eliminate all columns except transcript ID and TPM
TPMcounts <- TPMcounts %>%
  select(target_id, tpm)

# Rename columns for consistency and to ID TPM counts
colnames(TPMcounts)[1:2] <- c("Transcript_ID",
                              paste0("id", libraries[1], "_TPM"))

# Loop through remaining kallisto files, performing full joins to the kallisto file we read in
for (i in 2:length(kallisto_files)){
  idnum <- str_extract(kallisto_files[i], "id[0-9]+")
  kallisto_output <- read.delim(file = kallisto_files[i],
                                header = TRUE,
                                sep = "\t")
  # Select only transcript ID and TPM (transcripts per million) columns
  kallisto_output <- kallisto_output %>%
    select(target_id, tpm)
  # Rename kallisto column names to give ID to count column
  colnames(kallisto_output)[1:2] <- c("Transcript_ID", 
                                      paste0(idnum, "_TPM"))
  # Add TPM value to table of DEGs
  # Perform full join, keeping all transcript IDs
  TPMcounts <- full_join(TPMcounts, kallisto_output, by = "Transcript_ID")
}
```

WGCNA has several recommendations when it comes to RNAseq data, available in the FAQ [here](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/faq.html). First, they suggest removing all transcripts with counts below 10 in over 90% of samples. Since we have 15 samples total, we will only remove all transcripts with counts below 10 in 14 or 15 samples.

They also suggest a variance-stabilizing transformation or log-transforming the counts using log2(x+1). Unlike in our trial run of Crab A, we will be able to create a DESeq object and perform a variance-stabilizing transformation.

We will change our data to fit both of these recommendations. After, we will transpose the data frame so samples are rows and transcripts are columns

```{r rnaseqdataprep}
# Create logical matrix for whole dataframe, comparing values to 10

# Move transcript ID to rownames
TPMcounts <- TPMcounts %>%
  column_to_rownames(var = "Transcript_ID")

# Get initial dimensions of data frame
dim(TPMcounts)

# Get number of counts above 10 for each gene
TPMcounts$counts <- rowSums(TPMcounts >= 10)
# Remove all genes without multiple counts above 10
TPMcounts <- TPMcounts[TPMcounts$counts > 1, ]
# Remove the counts column we just created
TPMcounts <- subset(TPMcounts, select = - counts)

# See how many transcripts we have left
dim(TPMcounts)

# Round all counts to the nearest integer
TPMcounts <- round(TPMcounts, digits = 0)

# Normalize raw counts with DESeq()
crab.dds <- DESeqDataSetFromMatrix(countData = TPMcounts,
                                   colData = crabTraits,
                                   design = as.formula(paste0("~", variable)))
crab.dds <- DESeq(crab.dds)

# Perform vst on DESeq object
vsd <- getVarianceStabilizedData(crab.dds)

# Transpose dataframe to format for WGCNA
CrabExpr0 <- as.data.frame(t(vsd))

# Check dataframe was transposed correctly
dim(CrabExpr0)
```

We will now begin analysis with WGCNA. Our script is based largely on [Yaamini's WGCNA script](https://github.com/eimd-2019/project-EWD-transcriptomics/blob/master/analyses/WGCNA/WGCNA.md), which is based largely on [the WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/)

### Check for Missing Values and Outliers

```{r wgcna_missingvals}
# Check for genes and samples with too many missing values

gsg <- goodSamplesGenes(CrabExpr0, verbose = 3)
gsg$allOK      # should return TRUE if all genes pass test

sampleTree <- hclust(dist(CrabExpr0), method = "average")
```

```{r sample_tree}
path <- paste0(file_start, "ClusterDendrogram.png")
png(path)
plot(sampleTree)
dev.off()
# Plot image again, so it shows up in knitted .Rmd
plot(sampleTree)
```

### Create clinical trait data matrix

```{r datatraitmatrix}
# Print the crabTraits matrix we made earlier
head(crabTraits)
# Use same rownames as expression data to create analogous  matrix
rownames(crabTraits) <- rownames(CrabExpr0)
# Make sure it looks good
head(crabTraits)

# Create a dendrogram to look at sample and trait clustering
sampleTree2 <- hclust(dist(CrabExpr0), method = "average")
traitColors <- numbers2colors(crabClinicalData, signed = FALSE)
```

```{r sample_dendrogram}
path <- paste0(file_start, "ClusterDendrogram_W_Colors.png")
png(path)
plotDendroAndColors(sampleTree2, traitColors, 
                    groupLabels = names(crabTraits))
dev.off()
# Plot image again, so it shows up in knitted .Rmd
plotDendroAndColors(sampleTree2, traitColors, 
                    groupLabels = names(crabTraits))
```

### Network Connection and Module Detection

[Relevant tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/)

#### Determine soft-thresholding power

```{r softthresholding}
# Create set of soft-thresholding powers
powers <- c(c(1:10), seq(from = 12, to = 20, by = 2))
# Use network topology analysis function to eval soft-thresholding power vals
sft <- pickSoftThreshold(CrabExpr0, powerVector = powers, verbose = 5)
```

#### Plot scale-free topology fit as function of soft-thresholding power

```{r scale_free_topology_vs_soft_thresholding_power}
path <- paste0(file_start, "ScaleIndependence.png")
png(path)
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
                               labels = powers,
                               cex = 1,
                               col = "red")
abline(h=0.80, col = "black")
dev.off()

# Plot image again, so it shows up in knitted .Rmd
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
                               labels = powers,
                               cex = 1,
                               col = "red")
abline(h=0.80, col = "black")
```

#### Plot mean connectivity as function of soft-thresholding power

```{r mean_connectivity_vs_soft_thresholding_power}
path <- paste0(file_start, "MeanConnectivity.png")
png(path)
plot(sft$fitIndices[,1],sft$fitIndices[,5],
     xlab = "Soft Threshold (power)",
     ylab = "Mean Connectivity",
     type = "n",
     main = paste("Mean connectivity"))
# Add sft values
text(sft$fitIndices[,1], sft$fitIndices[,5],
     labels = powers,
     cex = 1,
     col = "red")
dev.off()

# Plot image again, so it shows up in knitted .Rmd
plot(sft$fitIndices[,1],sft$fitIndices[,5],
     xlab = "Soft Threshold (power)",
     ylab = "Mean Connectivity",
     type = "n",
     main = paste("Mean connectivity"))
# Add sft values
text(sft$fitIndices[,1], sft$fitIndices[,5],
     labels = powers,
     cex = 1,
     col = "red")
```

Typically, we would choose the lowest power below 15 that reached an R2 value of 0.8 or higher. However, no eligible soft-thresholding power reached 0.8. According to the [WGCNA FAQ](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/faq.html), this indicates the following:

"If the scale-free topology fit index fails to reach values above 0.8 for reasonable powers (less than 15 for unsigned or signed hybrid networks, and less than 30 for signed networks) and the mean connectivity remains relatively high (in the hundreds or above) [note: which our data does], chances are that the data exhibit a strong driver that makes a subset of the samples globally different from the rest. The difference causes high correlation among large groups of genes which invalidates the assumption of the scale-free topology approximation."

I chose to use an unsigned network, since the direction of correlation could be quite interesting (i.e. associating up-regulated C. bairdi with down-regulated Hemat.). Therefore, the WGCNA FAQ recommends choosing a soft-thresholding power of 9, since we have fewer than 20 samples total.


#### Co-expression similarity and adjacency, and Topological Overlap Matrix (TOM)

```{r coexpression}
softPower <- 9
adjacency <- adjacency(CrabExpr0, power = softPower)

# Minimize noise and spurious associations by transforming adjacency into TOM
TOM <- TOMsimilarity(adjacency)

#Calculate dissimilarity matrix
dissTOM <- 1 - TOM

# Clustering using TOM

# Create hierarchical clustering object
geneTree <- hclust(as.dist(dissTOM), method = "average")
```

#### Plot initial dendrogram. Dissimilarity is based on topological overlap
```{r gene_dendrogram}
path <- paste0(file_start, "GeneDendrogram.png")
png(path)
plot(geneTree, xlab = "", sub = "", 
     main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE,
     hang = 0.04)
dev.off()

# Plot image again, so it shows up in knitted .Rmd
plot(geneTree, xlab = "", sub = "", 
     main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE,
     hang = 0.04)
```

```{r min_module_size}
# Set minimum module size, AKA num of genes that need to be in a module. WGCNA default is 30
minModuleSize <- minimum_module
# Cut branches of dendrogram to ID WGCNA modules
dynamicMods <- cutreeDynamic(dendro =  geneTree,
                             distM = dissTOM,
                             deepSplit = 2,
                             pamRespectsDendro = FALSE,
                             minClusterSize = minModuleSize)

# Look at table of modules
table(dynamicMods)
# Convert module numbers into colors
dynamicColors <- labels2colors(dynamicMods)
```

```{r gene_dendrogram_module_colors}
path <- paste0(file_start, "GeneDendrogramWColors.png")
png(path)
# Plot dendrogram with module colors
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE,
                    hang = 0.03,
                    addGuide = TRUE,
                    guideHang = 0.05,
                    main = "Gene dendrogram and module colors")
dev.off()

# Plot image again, so it shows up in knitted .Rmd
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE,
                    hang = 0.03,
                    addGuide = TRUE,
                    guideHang = 0.05,
                    main = "Gene dendrogram and module colors")
```

#### Merge modules with similar expression profiles

Merging lets us combine modules with genes that are highly co-expressed. To do this, we can create and cluster eigengenes

```{r modulemerge}
# Calculate eigengenes
MElist <- moduleEigengenes(CrabExpr0, colors = dynamicColors)
# Save eigengenes as new object
MEs <- MElist$eigengenes
# Calculate dissimilarity of eigengenes
MEDiss <- 1-cor(MEs)
# Create cluster object
METree <- hclust(as.dist(MEDiss), method = "average")
```

```{r clustered_eigengenes}
path <- paste0(file_start, "ClusteredEigengenes.png")
png(path)
# Plot dendrogram of clustered eigengenes
plot(METree, main = "Clustering of module eigengenes",
     xlab = "",
     sub = "")
# ID cut height based on sample number
dynamicMergeCut(numsamples)

MEDissThres <- dynamicMergeCut(numsamples)
abline(h = MEDissThres, col = "red")
dev.off()

# Plot image again, so it shows up in knitted .Rmd
plot(METree, main = "Clustering of module eigengenes",
     xlab = "",
     sub = "")
# ID cut height based on sample number
dynamicMergeCut(numsamples)

MEDissThres <- dynamicMergeCut(numsamples)
abline(h = MEDissThres, col = "red")
```


```{r merge_close_modules}
merge <- mergeCloseModules(CrabExpr0, dynamicColors,
                           cutHeight = MEDissThres,
                           verbose = 3)

# Extract merged colors and eigengenes
mergedColors <- merge$colors
mergedMEs <- merge$newMEs
```

```{r dendrogram_w_orig_and_merged_eigengenes}
path <- paste0(file_start, "ClusterDendrogramOrigAndMergedEigengenes.png")
png(path)
# Plot dendrogram with original and merged eigengenes
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE,
                    hang = 0.03,
                    addGuide = TRUE,
                    guideHang = 0.05)
dev.off()

# Plot image again, so it shows up in knitted .Rmd
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"),
                    dendroLabels = FALSE,
                    hang = 0.03,
                    addGuide = TRUE,
                    guideHang = 0.05)
```

```{r save_merged_variables}
# Rename and save variables for subsequent analysis

moduleColors <- mergedColors
colorOrder <- c("grey", standardColors(50)) # Determine color order
moduleLabels <- match(moduleColors, colorOrder)-1 # Construct numerical labels based on colors
MEs <- mergedMEs # Replace unmerged MEs
```

### Relate modules to external traits

```{r module_trait_asscs}
# Count the number of genes and samples
nGenes <- ncol(CrabExpr0)
nSamples <- nrow(CrabExpr0)

# Recalculate MEs with color labels, order MEs based on MEs0
MEs0 <- moduleEigengenes(CrabExpr0, moduleColors)$eigengenes
MEs <- orderMEs(MEs0)

# Calculate trait correlations and obtain p-values
moduleTraitCor <- cor(MEs, crabClinicalData, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples)
moduleTraitPvalue

# Create text matrix for correlations and their p-values
textMatrix <- paste(signif(moduleTraitCor, 2), "\n(",
                        signif(moduleTraitPvalue, 1), ")", sep = "")
  
# Ensure matrix has same dimensions 
dim(textMatrix) == dim(moduleTraitCor)
```

```{r heatmap}
# Create labeled heat map of correlation values from textMatrix. Red = positive correlation, blue = negative correlation
path <- paste0(file_start, "ModuleTreatmentHeatMap.png")
png(path)
par(mar = c(4, 8.5, 3, 3))
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(crabTraits),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1, 1),
               main = paste("Module-Treatment relationships"))
dev.off()

# Plot image again, so it shows up in knitted .Rmd
par(mar = c(4, 8.5, 3, 3))
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(crabTraits),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1, 1),
               main = paste("Module-Treatment relationships"))
```

### Gene Significance and Module Membership

Module membership info is needed for a possible downstream GO-MWU analysis. 

``` {r gene_significance}
# Define "temp" using information from trait matrix
temp <- as.data.frame(crabClinicalData$temp)
# Modify names
names(temp) <- "temp"
# Save module names without "ME" at beginning of each entry
modNames <- substr(names(MEs), 3, nchar(names(MEs)))

# Determine gene significance

# Obtain gene significance statistics
geneTraitSignificance <- as.data.frame(cor(CrabExpr0, temp, use = "p"))
# Add column names
names(geneTraitSignificance) <- paste("GS.", names(temp), sep = "")
# Confirm formatting
head(geneTraitSignificance)

# Obtain p-values for each gene significance stat
GSPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples))
# Add column names
names(GSPvalue) <- paste("p.GS", names(temp), sep = "")
# Confirm formatting
head(GSPvalue)

# Determine module membership

# Obtain gene module membership stats
geneModuleMembership <- as.data.frame(cor(CrabExpr0, MEs, use = "p"))
# Add column names
names(geneModuleMembership) <- paste("MM", modNames, sep = "")
# Confirm formatting
head(geneModuleMembership)

# Obtain p-values for each module membership statistic
MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))
# Add column names
names(MMPvalue) <- paste("p.MM", modNames, sep = "")
# Confirm formatting
head(MMPvalue)
```

### Obtain gene lists - for each module, and a master list with membership and gene significance info

```{r genelists}
# Save gene names as probes
probes <- names(CrabExpr0)
# Write out the gene lists for each module of interest
for (module in modNames) {
  modGenes <- (moduleColors == module) # Select module probes
  modLLIDs <- probes[modGenes] # Get gene IDs
  fileName <- paste(file_start, "GeneList-", module, ".txt", sep = "") # Assign filename for each module
  write.table(as.data.frame(modLLIDs), file = fileName, sep = "\t", row.names = FALSE, col.names = FALSE) # Write out files
}
```

#### Master list with membership and gene significance information

```{r membership_and_gene_info}
# Import gene annotation info
crabGeneAnnot <- read.delim(blastx_table_site, header = FALSE, sep = "\t")
# Remove unnecessary columns
crabGeneAnnot <- crabGeneAnnot[, -c(3:10, 12)]
# Name columns
colnames(crabGeneAnnot) <- c("seqIDs", "Uniprot", "e-value")
# Look at column formatting
head(crabGeneAnnot)

# If pipes in Uniprot ID column, separate to specifically get Uniprot ID, and then remove those new columns with species info
gene_ids <- dplyr::pull(crabGeneAnnot, Uniprot)
if(any(grepl("|", gene_ids, fixed = TRUE))) {
  crabGeneAnnot <- separate(data = crabGeneAnnot, col = Uniprot, into = c("sp", "Uniprot", "Species"),
                   sep = "\\|")
  crabGeneAnnot <- crabGeneAnnot[,-c(2, 4)]
}

# Also get all GO terms
```