---
title: "62 - Mapping Pink Module"
author: "Aidan Coyle"
date: "Last compiled on `r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Earlier, we ran WGCNA on all ambient-temperature and elevated-temperature libraries over days 0, 2, and 17 (if applicable), tracking expression by day and by crab.

WGCNA was ran on libraries aligned to three transcriptomes - cbai_transcriptomev2.0 (unfiltered, includes both C. bairdi and Hematodinium), cbai_transcriptomev4.0 (BLASTed against C. opilio genome, should include only C. bairdi sequences), and hemat_transcriptomev1.6 (BLASTed against Alveolata sequences, should include only Hematodinium sequences)


Table of crabs and libraries included in analysis:

| Crab ID | Treatment Group | Day 0 Sample ID | Day 2 Sample ID | Day 17 Sample ID |
|---------|-----------------|-----------------|-----------------|------------------|
| A       | ambient         | 178             | 359             | 463              |
| B       | ambient         | 118             | 349             | 481              |
| C       | ambient         | 132             | 334             | 485  

And for elevated-temperature libraries:

| Crab ID | Treatment Group  | Day 0 Sample ID | Day 2 Sample ID |
|---------|------------------|-----------------|-----------------|
| G       | elevated         | 173             | 272             |
| H       | elevated         | 072             | 294             |
| I       | elevated         | 127             | 280             |

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(reshape2)
```

### Set up initial inputs to make script as easy to modify as possible

To prevent additional typing when switching from analyzing elevated to analyzing ambient, or vice-versa, here are the vectors for the following (can just copy-paste in the inputs block below). Be careful to also change inputs for modulepath!

``` {r vectors, eval = FALSE}
# Elevated Crabs

# Vector of all libraries included in module. If there are leading zeroes, put paste0("0", "[libnumber]") around those libraries
crablibs <- (c(173, 272, paste0("0", "72"), 294, 127, 280))

# Vector of variable of interest that corresponds to the libraries (should be in same order - if crablibs[3] is from Day 0, and your variable of interest is day, testedvar[3] should be 0)
testedvar <- c(0, 2, 0, 2, 0, 2)

crabids <- c("G", "G", "H", "H", "I", "I")

# Ambient Crabs

# Vector of all libraries included in module. If there are leading zeroes, put paste0("0", "[libnumber]") around those libraries
crablibs <- (c(178, 359, 463, 118, 349, 481, 132, 334, 485))

# Vector of variable of interest that corresponds to the libraries (should be in same order - if crablibs[3] is from Day 0, and your variable of interest is day, testedvar[3] should be 0)
testedvar <- c(0, 2, 17, 0, 2, 17, 0, 2, 17)

crabids <- c("A", "A", "A", "B", "B", "B", "C", "C", "C")

```


When 
```{r inputs}
# Path to directory containing modules
modulepath <- "../output/WGCNA_output/hemat_transcriptome_v1.6/amb_crabs_no_filter/"

# Specify location of kallisto libraries included in module
libpath <- "../output/kallisto_libraries/hemat_transcriptomev1.6/"

# Vector of all libraries included in module. If there are leading zeroes, put paste0("0", "[libnumber]") around those libraries
crablibs <- (c(178, 359, 463, 118, 349, 481, 132, 334, 485))

# Vector of variable of interest that corresponds to the libraries (should be in same order - if crablibs[3] is from Day 0, and your variable of interest is day, testedvar[3] should be 0)
testedvar <- c(0, 2, 17, 0, 2, 17, 0, 2, 17)

crabids <- c("A", "A", "A", "B", "B", "B", "C", "C", "C")

# Variable - this will be present only in the graphs
var <- "Day"
```

## Loop through module files in directory. For each, graph TPMs over time and log(TPM) over time

```{r load_data}
# Pull all files in folder containing modules
modulefiles <- list.files(modulepath)

# Select only files starting in GeneList
modulefiles <- str_subset(modulefiles, "^GeneList-[a-z]")

for (i in 1:length(modulefiles)) {
  # Produce path to specific module file
  moduleroute <- paste0(modulepath, modulefiles[i])
  
  # Each module filename will be something like GeneList-color.txt. Remove the GeneList- and .txt portions so we can write files more easily
  outpath <- str_remove(modulefiles[i], "GeneList-")
  outpath <- str_remove(outpath, ".txt")
  
  # Read in module data
  module <- read.delim(file = moduleroute, header = FALSE)
  
  # Rename first column of module data
  names(module)[1] <- "Transcript_ID"
  
  # Create inner loop. In each loop, add transcripts per million (tpm) for each crab library
  for (i in 1:length(crablibs)) { 
    # Make filepath to the crab library
    filepath <- paste0(libpath, "id", crablibs[i],
                       "/abundance.tsv")
    
    # Read in library
    crabdata <- read.delim(file = filepath, header = TRUE, sep = "\t")
    
    # Remove all columns except transcript ID and transcripts per million
    crabdata <- crabdata[,c(1,5)]
    
    # Rename both columns
    names(crabdata) <- c("Transcript_ID", paste0("id", crablibs[i]))
    
    # Left join TPM data to genes in module
    module <- left_join(module, crabdata, by = "Transcript_ID")
    
    # End of inner loop
    
    }
   
  # Move transcript IDs into rownames
  rownames(module) <- module$Transcript_ID
  module <- module[,-1]
  
  # Transpose dataframe
  module <- as.data.frame(t(module))
  
  # Add column with our testedvar input
  module$testedvar <- testedvar
  
  # Use melt to pivot the table longer
  module <- melt(module, id = "testedvar")
  
  # Add column with crab IDs
  module$crab <- rep(crabids)
  
  # Concatenate new column with transcript IDs and crab IDs so each crab-gene forms a single line over time
  module$variable <- str_c(module$crab, "_", module$variable)
  
  # Use ggplot to graph non-log-transformed counts. NOTE: since we're inside a for loop, we need to explicitly print ggplot
  path <- paste0(modulepath, outpath, "_module_TPM_graph_crab_colors_short_yaxis.png")
  png(path)
  print(ggplot() +
    geom_line(data = module, aes(x = testedvar,
                                 y = value,
                                 group = variable,
                                 color = crab),
              size = .001) +
    xlab(var) +
    ylab("Transcripts per million (TPM)")) +
    coord_cartesian(ylim = c(0,1500))
  dev.off()
  
  # Use ggplot to graph again, this time log transforming our TPM counts
  path <- paste0(modulepath, outpath, "_module_log_TPM_graph_crab_colors.png")
  png(path)
  print(ggplot() +
    geom_line(data = module, aes(x = testedvar,
                                 y = log(value),
                                 group = variable,
                                 color = crab),
              size = .001) +
    xlab(var) +
    ylab("Log-TPM"))
  dev.off()
  
  # Use ggplot to graph both again, just so it shows up in our knitted .Rmd file
  print(ggplot() +
    geom_line(data = module, aes(x = testedvar,
                                 y = value,
                                 group = variable,
                                 color = crab),
              size = .001) +
    xlab(var) +
    ylab("Transcripts per million (TPM)")) +
    ylim(0, 1500)    
  
  
  
  print(ggplot() +
    geom_line(data = module, aes(x = testedvar,
                                 y = log(value),
                                 group = variable,
                                 color = crab),
              size = .001) +
    xlab(var) +
    ylab("Log-TPM") +
    ylim(0, 1500))
}


```

