---
title: "7_7_merged_cluster_counts.Rmd"
author: "Aidan Coyle"
date: "7/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In scripts 7_1 through 7_3, we manually clustered gene expression from our 3 transcriptomes. Reminder: they are as follows:

cbai_transcriptomev2.0: unfiltered

cbai_transcriptomev4.0: filtered to include only likely _Chionoecetes_ sequences

hemat_transcriptomev1.6: filtered to include only likely _Alveolata_ sequences

However, once we named our modules, we had duplicate names. Modules describing the same expression patterns (as determined by names and assigned in 7_1 thru 7_3) were merged in scripts 7_4 through 7_6. At the end of each of those scripts, we wrote the word count - which is equal to the number of genes in each module - of the merged modules for each crab to a file. 

In this script, we will take those word counts and turn them into tables for optimal presentation and reproducibility


```{r libraries, message=FALSE, warning=FALSE}
# Add all required libraries here
list.of.packages <- c("tidyverse", "janitor", "ztable", "htmltools")
# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)


# Load all required libraries
lapply(list.of.packages, FUN = function(X) {
  do.call("require", list(X))
})

```

### Reading in data

```{r}
# Get the path of all relevant files
file_list <- Sys.glob("../output/manual_clustering/*/merged_modules_raw_counts.txt")

# In each iteration of the for loop, we'll choose a different transcriptome's raw counts to examine, create two neat summary tables - one with percentages, one with counts - and write as CSVs
for (i in 1:length(file_list)) {
  counts <- read.table(file_list[i])

  # Remove the last line - we can figure out the total on our own
  counts <- head(counts, -1)

  # Split the path column by slashes
  counts <- separate(counts, 2, into = c("A", "B", "C", "D", "E", "F", "G"), sep = "/")

  # Remove columns without multiple values. Should leave us with columns for counts, crab, and module type
  counts <- counts[vapply(counts, function(x) length(unique(x)) > 1, logical(1L))]
  
  # Rename existing columns
  colnames(counts) <- c("Genes", "Crab", "Module")
  
  # Remove the _merged.txt part of each Module column
  counts$Module <- str_replace(counts$Module, "_merged.txt", "")
  
  # Pivot wider so that each module type is its own column
  counts <- counts %>%
    pivot_wider(names_from = Module, values_from = Genes)
  
  # Create another table with percentage of module membership for each crab (each crab should sum to 100%)
  percentages <- adorn_percentages(counts, denominator = "row", na.rm = TRUE)
  
  # Move crab column to rowname for both tables
  counts <- column_to_rownames(counts, var = "Crab")
  percentages <- column_to_rownames(percentages, var = "Crab")
  
  percentages <- round(percentages, digits = 3)
  
  percs <- ztable(percentages)

  perc_table <- percs %>%
    makeHeatmap()
  
  print(perc_table)
  
  # Get the path for that transcriptome
  path <- file_list[i]
  
  # Remove the ending part of the path
  path <- str_replace(path, "merged_modules_raw_counts.txt", "")
  
  # Write our counts table
  write.csv(counts, file = paste0(path, "merged_modules_counts_table.csv"))
  
  # Save our percentages heatmap
  ggsave(filename = paste0(path, "merged_module_percentages_heatmap.png"),
         plot = perc_table, device = "png")
  
  # Open a PNG file and save our percentages heatmap
  png(file = paste0(path, "merged_module_percentages_heatmap.png"))
  print(makeHeatmap(percentages))
  dev.off()
  
  plot(cars$speed, cars$dist)
}

plot(cars)
counts <- read.table(file_list[1])

# Remove the last line - we can figure out the total on our own
counts <- head(counts, -1)

# Split the path column by slashes
counts <- separate(counts, 2, into = c("A", "B", "C", "D", "E", "F", "G"), sep = "/")

# Remove columns without multiple values. Should leave us with columns for counts, crab, and module type
counts <- counts[vapply(counts, function(x) length(unique(x)) > 1, logical(1L))]

# Rename existing columns
colnames(counts) <- c("Genes", "Crab", "Module")

# Remove the _merged.txt part of each Module column
counts$Module <- str_replace(counts$Module, "_merged.txt", "")

# Pivot wider so that each module type is its own column
counts <- counts %>%
  pivot_wider(names_from = Module, values_from = Genes)

# Create another table with percentage of module membership for each crab (each crab should sum to 100%)
percentages <- adorn_percentages(counts, denominator = "row", na.rm = TRUE)

# Move crab column to rowname for both tables
counts <- column_to_rownames(counts, var = "Crab")
percentages <- column_to_rownames(percentages, var = "Crab")

percentages <- round(percentages, digits = 3)




testper <- ztable(percentages)

testper %>%
  makeHeatmap() %>%
  print()

# Get the path for that transcriptome
path <- file_list[1]

# Remove the ending part of the path
path <- str_replace(path, "merged_modules_raw_counts.txt", "")

# Write our counts table
write.csv(counts, file = paste0(path, "merged_modules_counts_table.csv"))
# Write our percentages table
write.csv(percentages, file = paste0(path, "merged_module_percentages_table.csv"))

```

