---
title: "8_2_immune_PCAs"
author: "Aidan Coyle"
date: "8/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this script, we'll be taking the files of immune genes we previously created for each transcriptome, and using them to create PCAs, with each crab as an individual point.

### All Samples

The purpose of this section is to create a PCA of all libraries. Each point would be one library. This lets us visualize what variables are influencing variance.

Because of this, we'll actually run DESeq more directly rather than calling the deseq_analysis() function - no need for a lot of the typical DESeq2 outputs. That'll also let us use ggplotly() to create a more interactive (and thus easier to visualize) PCA.

After that, we'll also calculate the plasticity by getting the data from our PCA and calculating the mean distance between each point (using PC1/PC2) and the centroid of that crab.

We will not produce PCAs for hemat_transcriptomev1.6, as there were only 5 immune-linked genes for this transcriptome.

```{r libraries, message = FALSE, warning=FALSE}
# Add all required libraries that are installed with install.packages() here
list.of.packages <- c("tidyverse", "plotly")
# Add all libraries that are installed using BiocManager here
bioconductor.packages <- c("apeglm", "DESeq2", "vsn")

# Install BiocManager if needed
if(!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

# Get names of all required packages that aren't installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
new.bioc.packages <- bioconductor.packages[!(bioconductor.packages %in% installed.packages()[, "Package"])]
# Install all new packages
if(length(new.packages)) install.packages(new.packages)
if(length(new.bioc.packages)) BiocManager::install(new.bioc.packages)

# Load all required libraries
all.packages <- c(list.of.packages, bioconductor.packages)
lapply(all.packages, FUN = function(X) {
  do.call("require", list(X))
})
```

# cbai_transcriptomev2.0


```{r}
# Create experimental table. Order should correspond to order of columns in our data file
# In creating our initial data file, the easiest way to read the kallisto files was by pulling libraries in numeric order.
# Therefore, the order of both should be as follows:
# 072, 073, 113, 118, 127, 132, 151, 173, 178, 221, 222, 254, 272, 280, 294, 334, 349, 359, 425, 427, 445, 463, 481, 485

exp_design <- data.frame(crab = factor(c("H", "D", "F",
                                         "B", "I", "C",
                                         "E", "G", "A",
                                         "D", "F", "E",
                                         "G", "I", "H",
                                         "C", "B", "A",
                                         "F", "D", "E",
                                         "A", "B", "C")),
                         temp = factor(c("amb", "amb", "amb",
                                         "amb", "amb", "amb",
                                         "amb", "amb", "amb",
                                         "low", "low", "low",
                                         "elev", "elev", "elev",
                                         "amb", "amb", "amb",
                                         "low", "low", "low",
                                         "amb", "amb", "amb")),
                         day = factor(c(0, 0, 0,
                                        0, 0, 0,
                                        0, 0, 0,
                                        2, 2, 2,
                                        2, 2, 2, 
                                        2, 2, 2, 
                                        17, 17, 17,
                                        17, 17, 17)),
                         hemat_level = factor(c("H", "H", "L",
                                                "H", "H", "H",
                                                "L", "L", "L",
                                                "H", "L", "L",
                                                "L", "H", "H",
                                                "H", "H", "L",
                                                "L", "H", "L",
                                                "L", "H", "H")),
                         infection = factor(c("inf", "uninf", "uninf",
                                              "inf", "inf", "inf",
                                              "inf", "inf", "inf",
                                              "uninf", "uninf", "inf",
                                              "inf", "inf", "inf",
                                              "inf", "inf", "inf",
                                              "uninf", "uninf", "inf",
                                              "inf", "inf", "inf")))
# Get the names of each column in the exp_design table
variables <- colnames(exp_design)


for (i in 1:length(variables)) {
  # Select a variable from exp_design to show in your PCA
  var_name <- variables[i]
  # Read in data
  data <- read.table("../output/kallisto_matrices/cbai_transcriptomev2.0/immune_genes_indiv_libraries/manual_cts_matrix.txt", header = TRUE, sep = "\t")
  
  # Rename first column
  names(data)[1] <- "target_ID"
  
  # Set row names equal to the first column
  rownames(data) <- data$target_ID
  
  # Remove the now-irrelevant first column
  data <- data[,-1]
  
  # Make sure everything looks okay
  print("HEAD")
  head(data)
  
  print("STRUCTURE")
  print(str(data))
  
  # Round counts to integers - needed for DESeqDataSetFromMatrix()
  data <- round(data, digits = 0)
  
  # Rename rows to correspond to library numbers
  rownames(exp_design) <- colnames(data)
  
  # Check that experiment_table appears to match columns with matrix from Trinity
  print("EXPERIMENTAL DESIGN")
  print(exp_design)
  
  # Create DESeq object that looks at effect of variable
  deseq2.dds <- DESeqDataSetFromMatrix(countData = (data),
                                       colData = exp_design,
                                       design = as.formula("~temp"))
  
  deseq2.dds <- DESeq(deseq2.dds)
  
  #Look at results
  deseq2.res <- results(deseq2.dds)
  print("SUMMARY:")
  print(summary(deseq2.res))
  
  # Transform values for PCA plot
  vsd <- varianceStabilizingTransformation(deseq2.dds, blind = FALSE)
  
  colData(vsd)
  
  # Plot plotly for PCA
  png(paste0("../graphs/DESeq2_output/cbai_transcriptomev2.0/immune_genes_all_libs/",
             var_name, "_as_var.png"))
  
  print(plotPCA(vsd, intgroup= var_name) +
    ggtitle("PCA of all crabs for cbai_transcriptomev2.0") +
    geom_point(size = 3, aes(text = colnames(vsd))) +
    theme_minimal())
  dev.off()
}

# Now that we're outside the for loop, we can calculate the centroid of the PCA for each crab to get an idea of the plasticity. To do this, we'll take the mean distance between each library and the centroid.

# Run PCA to create object. This uses the same code found in plotPCA()
rv <- rowVars(assay(vsd))
select <- order(rv, decreasing = TRUE)[seq_len(min(500, 
                                                   length(rv)))]
pca <- prcomp(t(assay(vsd)[select, ]))

# Find the points of each PCA
points <- pca[["x"]]

# Change rownames from sample ID to crab ID
rownames(points) <- exp_design$crab

# Eliminate all columns except PC1 and PC2
points <- points[,1:2]

# Get list of all rownames used (should be all unique crab IDs)
crabs <- unique(rownames(points))

# Initialize a blank dataframe
means <- data.frame(crabs, rep(NA, times = length(crabs)), rep(NA, times = length(crabs)), rep(NA, times = length(crabs)))

# Rename the columns of the blank dataframe
colnames(means) <- c("crabID", "PC1_mean", "PC2_mean", "mean_dist_from_centroid")

for (i in 1:length(crabs)) {
  # Get values that match that particular crab
  crab_pts <- points[rownames(points) == crabs[i], ]
  
  # Get mean of each column. This is the coordinates of the centroid
  means[i, 2] <- mean(crab_pts[, 1])
  means[i, 3] <- mean(crab_pts[, 2])
  
  # Initialize blank vector that's the same length as the number of values that match the crab in question
  dist_vec <- rep(NA, length = nrow(crab_pts))
  
  for (j in 1:nrow(crab_pts)) {
    # Calculate the Euclidean distance from each point to the centroid, use to fill the blank vector
    dist_vec[j] <- sqrt((crab_pts[j,1] - means[i, 2])^2 + (crab_pts[j,2] - means[i, 3])^2)
    
  }
  # Take the mean of the Euclidean distances and put that in Column 4 of the means table
  means[i,4] <- mean(dist_vec)
}

# Remove the middle 2 columns, since we really just want mean distance in the final table
means <- select(means, -c(PC1_mean, PC2_mean))


write.table(means, file = "../graphs/DESeq2_output/cbai_transcriptomev2.0/immune_genes_all_libs/PCA_plasticity.txt", row.names = FALSE)

```

# cbai_transcriptomev4.0


```{r}
# Create experimental table. Order should correspond to order of columns in our data file
# In creating our initial data file, the easiest way to read the kallisto files was by pulling libraries in numeric order.
# Therefore, the order of both should be as follows:
# 072, 073, 113, 118, 127, 132, 151, 173, 178, 221, 222, 254, 272, 280, 294, 334, 349, 359, 425, 427, 445, 463, 481, 485

exp_design <- data.frame(crab = factor(c("H", "D", "F",
                                         "B", "I", "C",
                                         "E", "G", "A",
                                         "D", "F", "E",
                                         "G", "I", "H",
                                         "C", "B", "A",
                                         "F", "D", "E",
                                         "A", "B", "C")),
                         temp = factor(c("amb", "amb", "amb",
                                         "amb", "amb", "amb",
                                         "amb", "amb", "amb",
                                         "low", "low", "low",
                                         "elev", "elev", "elev",
                                         "amb", "amb", "amb",
                                         "low", "low", "low",
                                         "amb", "amb", "amb")),
                         day = factor(c(0, 0, 0,
                                        0, 0, 0,
                                        0, 0, 0,
                                        2, 2, 2,
                                        2, 2, 2, 
                                        2, 2, 2, 
                                        17, 17, 17,
                                        17, 17, 17)),
                         hemat_level = factor(c("H", "H", "L",
                                                "H", "H", "H",
                                                "L", "L", "L",
                                                "H", "L", "L",
                                                "L", "H", "H",
                                                "H", "H", "L",
                                                "L", "H", "L",
                                                "L", "H", "H")),
                         infection = factor(c("inf", "uninf", "uninf",
                                              "inf", "inf", "inf",
                                              "inf", "inf", "inf",
                                              "uninf", "uninf", "inf",
                                              "inf", "inf", "inf",
                                              "inf", "inf", "inf",
                                              "uninf", "uninf", "inf",
                                              "inf", "inf", "inf")))

# Get the names of each column in the exp_design table
variables <- colnames(exp_design)


for (i in 1:length(variables)) {
  # Select a variable from exp_design to show in your PCA
  var_name <- variables[i]
  
  # Read in data
  data <- read.table("../output/kallisto_matrices/cbai_transcriptomev4.0/immune_genes_indiv_libraries/manual_cts_matrix.txt", header = TRUE, sep = "\t")
  
  # Rename first column
  names(data)[1] <- "target_ID"
  
  # Set row names equal to the first column
  rownames(data) <- data$target_ID
  
  # Remove the now-irrelevant first column
  data <- data[,-1]
  
  # Make sure everything looks okay
  print("HEAD")
  head(data)
  
  print("STRUCTURE")
  print(str(data))
  
  # Round counts to integers - needed for DESeqDataSetFromMatrix()
  data <- round(data, digits = 0)
  
  # Rename rows to correspond to library numbers
  rownames(exp_design) <- colnames(data)
  
  # Check that experiment_table appears to match columns with matrix from Trinity
  print("EXPERIMENTAL DESIGN")
  print(exp_design)
  
  # Create DESeq object that looks at effect of variable
  deseq2.dds <- DESeqDataSetFromMatrix(countData = (data),
                                       colData = exp_design,
                                       design = as.formula("~temp"))
  
  deseq2.dds <- DESeq(deseq2.dds)
  
  #Look at results
  deseq2.res <- results(deseq2.dds)
  print("SUMMARY:")
  print(summary(deseq2.res))
  
  # Transform values for PCA plot
  vsd <- varianceStabilizingTransformation(deseq2.dds, blind = FALSE)
  
  colData(vsd)
  
  # Plot plotly for PCA
  png(paste0("../graphs/DESeq2_output/cbai_transcriptomev4.0/immune_genes_all_libs/",
             var_name, "_as_var.png"))
  
  print(plotPCA(vsd, intgroup= var_name) +
          ggtitle("PCA of all crabs for cbai_transcriptomev4.0") +
          geom_point(size = 3, aes(text = colnames(vsd))) +
          theme_minimal())
  dev.off()
}

# Now that we're outside the for loop, we can calculate the centroid of the PCA for each crab to get an idea of the plasticity. To do this, we'll take the mean distance between each library and the centroid.

# Run PCA to create object. This uses the same code found in plotPCA()
rv <- rowVars(assay(vsd))
select <- order(rv, decreasing = TRUE)[seq_len(min(500, 
                                                   length(rv)))]
pca <- prcomp(t(assay(vsd)[select, ]))

# Find the points of each PCA
points <- pca[["x"]]

# Change rownames from sample ID to crab ID
rownames(points) <- exp_design$crab

# Eliminate all columns except PC1 and PC2
points <- points[,1:2]

# Get list of all rownames used (should be all unique crab IDs)
crabs <- unique(rownames(points))

# Initialize a blank dataframe
means <- data.frame(crabs, rep(NA, times = length(crabs)), rep(NA, times = length(crabs)), rep(NA, times = length(crabs)))

# Rename the columns of the blank dataframe
colnames(means) <- c("crabID", "PC1_mean", "PC2_mean", "mean_dist_from_centroid")

for (i in 1:length(crabs)) {
  # Get values that match that particular crab
  crab_pts <- points[rownames(points) == crabs[i], ]
  
  # Get mean of each column. This is the coordinates of the centroid
  means[i, 2] <- mean(crab_pts[, 1])
  means[i, 3] <- mean(crab_pts[, 2])
  
  # Initialize blank vector that's the same length as the number of values that match the crab in question
  dist_vec <- rep(NA, length = nrow(crab_pts))
  
  for (j in 1:nrow(crab_pts)) {
    # Calculate the Euclidean distance from each point to the centroid, use to fill the blank vector
    dist_vec[j] <- sqrt((crab_pts[j,1] - means[i, 2])^2 + (crab_pts[j,2] - means[i, 3])^2)
    
  }
  # Take the mean of the Euclidean distances and put that in Column 4 of the means table
  means[i,4] <- mean(dist_vec)
}

# Remove the middle 2 columns, since we really just want mean distance in the final table
means <- select(means, -c(PC1_mean, PC2_mean))


write.table(means, file = "../graphs/DESeq2_output/cbai_transcriptomev4.0/immune_genes_all_libs/PCA_plasticity.txt", row.names = FALSE)


```




The chunk below will find the plasticity of each crab by taking the mean distance from each point to the centroid of the PCA, and thus only needs to be run once.

```{r}
# Run PCA to create object. This uses the same code found in plotPCA()
rv <- rowVars(assay(vsd))
select <- order(rv, decreasing = TRUE)[seq_len(min(500, 
                                                   length(rv)))]
pca <- prcomp(t(assay(vsd)[select, ]))

# Find the points of each PCA
points <- pca[["x"]]

# Change rownames from sample ID to crab ID
rownames(points) <- exp_design$crab

# Eliminate all columns except PC1 and PC2
points <- points[,1:2]

# Get list of all rownames used (should be all unique crab IDs)
crabs <- unique(rownames(points))

# Initialize a blank dataframe
means <- data.frame(crabs, rep(NA, times = length(crabs)), rep(NA, times = length(crabs)), rep(NA, times = length(crabs)))

# Rename the columns of the blank dataframe
colnames(means) <- c("crabID", "PC1_mean", "PC2_mean", "mean_dist_from_centroid")

for (i in 1:length(crabs)) {
  # Get values that match that particular crab
  crab_pts <- points[rownames(points) == crabs[i], ]
  
  # Get mean of each column. This is the coordinates of the centroid
  means[i, 2] <- mean(crab_pts[, 1])
  means[i, 3] <- mean(crab_pts[, 2])
  
  # Initialize blank vector that's the same length as the number of values that match the crab in question
  dist_vec <- rep(NA, length = nrow(crab_pts))
  
  for (j in 1:nrow(crab_pts)) {
    # Calculate the Euclidean distance from each point to the centroid, use to fill the blank vector
    dist_vec[j] <- sqrt((crab_pts[j,1] - means[i, 2])^2 + (crab_pts[j,2] - means[i, 3])^2)
    
  }
  # Take the mean of the Euclidean distances and put that in Column 4 of the means table
  means[i,4] <- mean(dist_vec)
}

# Remove the middle 2 columns, since we really just want mean distance in the final table
means <- select(means, -c(PC1_mean, PC2_mean))


write.table(means, file = "../graphs/DESeq2_output/cbai_transcriptomev4.0/immune_genes_all_libs/PCA_plasticity.txt", row.names = FALSE)

```

