---
title: "Sample Summary"
author: "Aidan Coyle"
date: "2021-11-01"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include = FALSE}
library(tidyverse)
```


## Introduction

Samples were shipped from Pam Jensen to the Roberts Lab upon her retirement from NOAA. 

Uses Pam's data tables - which have not yet been cross-referenced with the physical samples - as the data source.

Total samples: 27,387


## Background: Sample Storage

- Majority: hemolymph samples in deep well plates
- Minority: tissue samples in tubes
- All: Preserved in 100% EtOH

## Background: Sample Location

```{r}
crabdat <- read.csv("../output/jensen_data/cleaned_data/relevant_sample_data.csv")

filt_crabdat <- crabdat

filt_crabdat$General_Location <- recode(filt_crabdat$General_Location, 
                                   "BS" = "Bering Sea",
                                   "GL" = "Greenland",
                                   "GOA" = "Gulf of Alaska",
                                   "OC" = "Oregon Coast",
                                   "SE" = "Southeast AK")

filt_crabdat %>%
  ggplot(aes(x = General_Location)) +
  geom_bar(fill = "steelblue", color = "steelblue") + 
  labs(x = "General Area", 
       y = "Number of Samples") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```

## Background: Sample Taxa

```{r}
filt_crabdat <- crabdat
filt_crabdat$Species_Name <- filt_crabdat$Species_Name

filt_crabdat$Species_Name <- recode(filt_crabdat$Species_Name, 
                                    "Cancer magister" = "Cancer sp.",
                                    "Cancer oregonensis" = "Cancer sp.",
                                    "Cancer productus" = "Cancer sp.",
                                    "Chionoecetes bairdi" = "C. bairdi",
                                    "Chionoecetes opilio" = "C. opilio",
                                    "Hyas coarctatus" = "Hyas sp.",
                                    "Hyas lyratus" = "Hyas sp.",
                                    "Labidochirus splendescens" = "Pagurid",
                                    "Lithodes aequispinus" = "Lithodid",
                                    "Pagurus aleuticus" = "Pagurid",
                                    "Pagurus confragosus" = "Pagurid",
                                    "Pagurus ochotensis" = "Pagurid",
                                    "Pagurus rathbuni" = "Pagurid",
                                    "Pagurus sp." = "Pagurid",
                                    "Pagurus trigonocheirus" = "Pagurid",
                                    "Pagurus platyceros " = "Pagurid",
                                    "Paralithodes camtschaticus" = "Lithodid",
                                    "Paralithodes platypus" = "Lithodid",
                                    "Pandalus platyceros" = "other",
                                    "Placetron wosnessenskii" = "Lithodid",
                                    "Telmessus cheiragonus" = "other", 
                                    "Elassochirus cavimanus" = "Pagurid",
                                    "Erimacrus isenbeckii" = "other",
                                    "Chionoecetes hybrid" = "other")

filt_crabdat %>%
  ggplot(aes(x = as.factor(Species_Name))) +
  geom_bar(fill = "steelblue", color = "steelblue") + 
  labs(x = "Taxa", 
       y = "Number of Samples") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
```

## Samples of Interest: Bering Sea 2018-2019

Graph of positive Hematodinium rates by year

## Samples of Interest: Bering Sea 2018-2019

Summary of samples by species and location

## Lab Resources

Summarize lab resources, approx. cost to sequence (+ cost of labor?)

## 




## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

